Farley et al. 2024 B

### 2025-01-12 IndividualStudies/SF_2024_B.qmd
Holy cow my git seems confused

---
title: "Farley et al. 2024"
subtitle: "Trifunctional Sphinganine: A New Tool to Dissect Sphingolipid Function"
format: 
  html:
    embed-resources: false
standalone: false
---

@@

## This is a little messy, but converts the raw data Scotty received from Frank and Per to a clean and minimal dataframe.
SF_2024_data <- SF_2024_data_raw |>
	select(gene_name, keep_Huh7, comparison, logFC, pvalue, hit_annotation) |>
	filter(keep_Huh7 == TRUE) |>
	select(-keep_Huh7) |>
	mutate(LipidProbe = case_when(comparison == "FA_8_3_UV - FA_8_3_noUV" ~ "8-3 FA",
															 comparison == "FA_1_10_UV - FA_1_10_noUV" ~ "1-10 FA",
															 comparison == "Spa_UV - Spa_noUV" ~ "Spa",
															 comparison == "Sph_UV - Sph_noUV" ~ "Sph")) |>
	filter(LipidProbe == "8-3 FA" | LipidProbe == "1-10 FA" | LipidProbe == "Spa" | LipidProbe == "Sph") |>
	select(-comparison) |>
	glimpse()

@@

# Building an ID column to maintain the order on the x-axis
ID <- c()
for(i in 0:(nrow(SF_2024_data)/4)){
	ID[i] = i
}

@@

# Applies the ID column to the Spa results
SF_2024_ranked_Spa <- SF_2024_data |>
    arrange(logFC) |>
	filter(LipidProbe == "Spa") |>
	mutate(ID = ID)

# Applies the ID column to the Sph results
SF_2024_ranked_Sph <- SF_2024_data |>
    arrange(logFC) |>
	filter(LipidProbe == "Sph") |>
	mutate(ID = ID)

# Assembles the dataset to plot
SF_2024_ranked <- rbind(SF_2024_ranked_Sph, SF_2024_ranked_Spa, SF_2024_ranked_8_3_FA, SF_2024_ranked_1_10_FA)
```

Scotland E. Farley, Frank Stein, Per Haberkant, Fikadu G. Tafesse, Carsten Schultz*

@@

2024-02-16

[ACS Chemical Biology](https://pubs.acs.org/doi/10.1021/acschembio.3c00554){target="_blank"}

@@

Functions of the sphingolipids sphingosine and sphinganine in cells are not well established. While some signaling roles for sphingosine have been elucidated, the closely related sphinganine has been described only insofar as it does not elicit many of the same signaling responses. The underlying mechanisms behind the cell biological differences between these lipids are not well understood. Here, we prepared multifunctionalized derivatives of the two lipid species that only differ in a single double bond of the carbon backbone. Using these novel probes, we were able to define their spatiotemporal distribution within cells. Furthermore, we used these tools to systematically map the protein interactomes of both lipids. The lipid-protein conjugates, prepared through photo-crosslinking in live cells and extraction via click chemistry to azide beads, revealed significant differences in the captured proteins, highlighting their distinct roles in various cellular processes. This work elucidates mechanistic differences between these critical lipids and sets the foundation for further studies on the functions of sphingosine and sphinganine.

@@

{{< downloadthis DataTables/SF_ACS-Chem-Bio_2024_download.csv dname="SF_ACS-Chem-Bio_2024">}}

>>>>

---
title: "Farley et al. 2024 B"
subtitle: "Trifunctional fatty acid derivatives: the impact of diazirine placement"
format: 
  html:
    embed-resources: false
standalone: false
---

@@

## This is a little messy, but converts the raw data Scotty received from Frank and Per to a clean and minimal dataframe.
SF_2024_data <- SF_2024_data_raw |>
	select(gene_name, keep_Huh7, comparison, logFC, pvalue, hit_annotation) |>
	filter(keep_Huh7 == TRUE) |>
	select(-keep_Huh7) |>
	mutate(LipidProbe = case_when(comparison == "FA_8_3_UV - FA_8_3_noUV" ~ "8-3 FA",
															 comparison == "FA_1_10_UV - FA_1_10_noUV" ~ "1-10 FA")) |>
	filter(LipidProbe == "8-3 FA" | LipidProbe == "1-10 FA" | LipidProbe == "Spa" | LipidProbe == "Sph") |>
	select(-comparison) |>
	glimpse()

@@

#write.csv(SF_2024_data, "IndividualStudies/DataTables/SF_Chem_Comm_2024_download.csv")

# Building an ID column to maintain the order on the x-axis
ID <- c()
for(i in 0:(nrow(SF_2024_data)/2)){
	ID[i] = i
}

@@

# Assembles the dataset to plot
SF_2024_ranked <- rbind(SF_2024_ranked_8_3_FA, SF_2024_ranked_1_10_FA)
```

Scotland E. Farley, Ryu Hashimoto, Judah Evangelista, Frank Stein, Per Haberkant, Kazuya Kikuchi, Fikadu G. Tafesse, Carsten Schultz*

@@

2024-05-25

[Chemical Communications](https://xlink.rsc.org/?DOI=D4CC00974F){target="_blank"}

@@

Functionalized lipid probes are a critical new tool to interrogate the function of individual lipid species, but the structural parameters that constrain their utility have not been thoroughly described. Functionalized lipid probes are a critical new tool to interrogate the function of individual lipid species, but the structural parameters that constrain their utility have not been thoroughly described. Here, we synthesize three palmitic acid derivatives with a diazirine at different positions on the acyl chain and examine their metabolism, subcellular localization, and protein interactions. We demonstrate that while they produce very similar metabolites and subcellular distributions, probes with the diazirine at either end pulldown distinct subsets of proteins after photo-crosslinking. This highlights the importance of thoughtful diazirine placement when developing probes based on biological molecules.

@@

{{< downloadthis DataTables/SF_Chem_Comm_2024_download.csv dname="SF_Chem_Comm_2024">}}
